Classification of ordinal data in deep learning: an experimental study

### 2019-03-15 paper/ieee.tex
Correcciones paper y nuevas funciones generalizadas

	Ordinal classification problems are those classification tasks where labels are ordered, and there are different inter-classes importances for each pair of classes. This kind of problem can be treated as a nominal classification problem, but this discards ordinal information. A better approach is to use specific methods that take into account this kind of information to improve the performance of the classification model. The Proportional Odds Model (POM)~\cite{agresti2010analysis} is an ordinal alternative to the softmax function that was designed for ordinal regression. It belongs to a wider family of models called Cumulative Link Models (CLM). It is inspired in the concept of a latent variable that is projected in an n-dimensional space and a set of thresholds that divides this space into the different ordinal levels. This kind of models uses a link function which can be of different types. The most common link function is the \texttt{logit} one, that is used in POM. However, there are other functions that can be explored. This model will be described in depth in Section \ref{sect:ordinalproblem}.

@@

	In this paper, an experimental study regarding link functions for Cumulative Link Models will be made. Also, other parameters that can affect the training process and the model performance, like the learning rate of the optimization algorithm and the batch size, and its interaction will be studied. The nominal version of this model will be used as a baseline for comparison. We will contrast the results obtained with statistical analysis to provide more robust conclusions. An approximated ANOVA III \cite{miller1997beyond} test followed by a posthoc Tukey's test \cite{tukey1949comparing} will be performed because of the limitations of the computational time required to run a higher number of executions.

@@

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning. First, they defined QWK metric as follows:
	\begin{equation}
	\text{QWK} = 1 - \frac{\sum\limits_{i,j} \omega_{i,j} O_{i,j}}{\sum\limits_{i,j} \omega_{i,j} E_{i,j}},
	\end{equation}
	where $\omega$ is the penalization matrix (in this case, quadratic weights are considered), $O$ is the confusion matrix and $E$ is the normalized outer product between the prediction and the true vector.

@@

		\begin{figure}[!t]
			\centering
			\begin{tabular}{cccc}
				\includegraphics[width=1.6cm]{img/adience/1.jpg} & \includegraphics[width=1.6cm]{img/adience/2.jpg} & \includegraphics[width=1.6cm]{img/adience/3.jpg} & \includegraphics[width=1.6cm]{img/adience/4.jpg}\\
				\includegraphics[width=1.6cm]{img/adience/5.jpg} & \includegraphics[width=1.6cm]{img/adience/6.jpg} & \includegraphics[width=1.6cm]{img/adience/7.jpg} & \includegraphics[width=1.6cm]{img/adience/8.jpg}
			\end{tabular}						
			\caption{Examples of the Diabetic Retinopathy test set.}
			\label{fig:AdienceExamples}
		\end{figure}
	\end{itemize}

@@

	Fig. \ref{fig:lrdecay} represents the learning rate decay for the different initial values considered in this work.
	\begin{figure}[!t]
		\centering
		\includegraphics[width=3in]{img/lr.pdf}
		\caption{Representation of the learning rate decay.}
		\label{fig:lrdecay}
	\end{figure}

>>>>

	Ordinal classification problems are those classification tasks where labels are ordered, and there are different inter-classes importances for each pair of classes. This kind of problem can be treated as a nominal classification problem, but doing this discards the ordinal information. A better approach is to use specific methods that take into account this kind of information to improve the performance of the classification model. The Proportional Odds Model (POM)~\cite{agresti2010analysis} is an ordinal alternative to the softmax function that was designed for ordinal regression. It belongs to a wider family of models called Cumulative Link Models (CLM). It is inspired in the concept of a latent variable that is projected in an n-dimensional space and a set of thresholds that divides that space into the different ordinal levels. This kind of models uses a link function which can be of different types. The most common link function is the \texttt{logit} one, that is used in POM. However, there are other functions that can be explored. This model will be described in depth in Section \ref{sect:ordinalproblem}.

@@

	In this paper, an experimental study regarding link functions for Cumulative Link Models in deep learning will be made. Also, other parameters that can affect the training process and the model performance, like the learning rate of the optimization algorithm, the batch size, and its interaction will be studied. The nominal version of this model will be used as a baseline for comparison. We will contrast the results obtained with statistical analysis to provide more robust conclusions. An approximated ANOVA III \cite{miller1997beyond} test followed by a posthoc Tukey's test \cite{tukey1949comparing} will be performed because of the limitations of the computational time required to run a higher number of executions.

@@

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK metric as loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning. First, they defined QWK metric as follows:
	\begin{equation}
	\text{QWK} = 1 - \frac{\sum\limits_{i,j} \omega_{i,j} O_{i,j}}{\sum\limits_{i,j} \omega_{i,j} E_{i,j}},
	\end{equation}
	where $\omega$ is the penalization matrix (in this case, quadratic weights are considered), $O$ is the confusion matrix and $E$ is the normalized outer product between the prediction and the true vector.

@@

		\begin{figure}[!t]
			\centering
			\begin{tabular}{cccc}
				\includegraphics[width=1.6cm]{img/adience/1.jpg} & \includegraphics[width=1.6cm]{img/adience/2.jpg} & \includegraphics[width=1.6cm]{img/adience/3.jpg} & \includegraphics[width=1.6cm]{img/adience/4.jpg}\\
				\includegraphics[width=1.6cm]{img/adience/5.jpg} & \includegraphics[width=1.6cm]{img/adience/6.jpg} & \includegraphics[width=1.6cm]{img/adience/7.jpg} & \includegraphics[width=1.6cm]{img/adience/8.jpg}
			\end{tabular}						
			\caption{Examples of the Adience test set.}
			\label{fig:AdienceExamples}
		\end{figure}
	\end{itemize}

@@

	Fig. \ref{fig:lrdecay} represents the learning rate decay across $100$ epochs for the different initial values considered in this work.
	\begin{figure}[!t]
		\centering
		\includegraphics[width=3in]{img/lr.pdf}
		\caption{Representation of the learning rate decay.}
		\label{fig:lrdecay}
	\end{figure}
